---
title: "Bayesian Spring Seminar 2019 - Week 1"
subtitle: "⚔<br/>Group Project"
author: "Gijs, Brayton & Fabio"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    toc: true
---

## Packages and Folders

```{r}
# Install these packages if you don't have them yet
# if (!require("pacman")) install.packages("pacman")
# devtools::install_github("favstats/tidytemplate")

pacman::p_load(tidyverse, tidytemplate, dummies, janitor, sjmisc, rjags, haven, ggthemes, ggmcmc, texreg, gridExtra, tidybayes, sjPlot, broom)


knitr::opts_chunk$set(fig.width = 8, warning = F, message = F, echo = T)

```


## Load and Clean Data

```{r, data}
## set seed first so it's reproducible!
set.seed(14042019)

class_dat <- read_sav("data/test Baysesian v2.sav") %>% 
  clean_names() %>% 
  mutate(volunteer = ifelse(v47 == 2, 0, v47)) %>% 
  mutate(social_class = sjmisc::to_label(social_class)) %>% 
  mutate(id = 1:n())  

class_dat <- class_dat %>% 
  to_dummy(social_class, suffix = "label") %>% 
  bind_cols(class_dat, .) %>% 
  clean_names() %>% 
  rename_all(~str_remove(.x, "social_class_"))

```

### Sample Description

```{r}
 class_dat %>% 
  count(social_class) %>% 
  ggplot(aes(social_class, n)) +
  geom_col(alpha = 0.5, fill = "grey") +
  coord_flip() +
  geom_text(aes(label = n)) +
  theme_minimal()


 class_dat %>% 
   drop_na(volunteer) %>% 
   count(social_class, volunteer) %>% 
   group_by(social_class) %>% 
   mutate(total = sum(n)) %>% 
   mutate(perc = n / total) %>% 
   filter(volunteer == 1) %>% 
   ggplot(aes(social_class, perc)) +
   geom_col(fill = "grey") +
   coord_flip() +
   geom_text(aes(label = round(perc*100))) +
   theme_minimal()

```

## Draw sample

```{r}

poor_sample <- class_dat %>% 
  sample_n(500)

rich_sample <- class_dat %>% 
  anti_join(poor_sample %>% select(id))
  
```

### Poor Sample Description

```{r}
 poor_sample %>% 
  count(social_class) %>% 
  ggplot(aes(social_class, n)) +
  geom_col(alpa = 0.5, fill = "grey") +
  coord_flip() +
  geom_text(aes(label = n)) +
  theme_minimal()


 class_dat %>% 
   drop_na(volunteer) %>% 
   count(social_class, volunteer) %>% 
   group_by(social_class) %>% 
   mutate(total = sum(n)) %>% 
   mutate(perc = n / total) %>% 
   filter(volunteer == 1) %>% 
   ggplot(aes(social_class, perc)) +
   geom_col(alpa = 0.5, fill = "grey") +
   coord_flip() +
   geom_text(aes(label = round(perc*100))) +
   theme_minimal()

```

## Rich People Analysis

Gezellig et al. 2019

```{r, fig.width=10}
rich_sample <- rich_sample %>% 
  ## Make Secure middle class to reference category
  mutate(social_class = fct_relevel(social_class, "Secure middle class"))

## fit Logistic Regression
glm_mod <- glm(volunteer ~ social_class, data = rich_sample, family = "binomial")

summary(glm_mod)
```

```{r, results='asis'}
texreg::htmlreg(glm_mod)
```


```{r}
sjPlot::plot_model(glm_mod, show.values = T, show.p = T)

glm_plot <- glm_mod %>% sjPlot::plot_model(type = "pred", show.intercept = T, show.values = T) 

glm_plot$social_class + coord_flip()

save_it(glm_mod)
```


## Flat Priors

### Model Specification

```{r binary_model2, echo = T}
flat_prior_mod <- "model{
  for (i in 1:N){
    y[i] ~ dbern(p[i])
    logit(p[i]) <- alpha + beta1 * x1[i] + beta2 * x2[i]+ beta3 * x3[i] + beta4 * x4[i]+ beta5 * x5[i] + beta6 * x6[i] + beta7 * x7[i] 
  }
  alpha ~ dnorm(0, 0.0001)
  beta1 ~ dnorm(0, 0.0001)
  beta2 ~ dnorm(0, 0.0001)
  beta3 ~ dnorm(0, 0.0001)
  beta4 ~ dnorm(0, 0.0001)
  beta5 ~ dnorm(0, 0.0001)
  beta6 ~ dnorm(0, 0.0001)
  beta7 ~ dnorm(0, 0.0001)
}"

write(flat_prior_mod, "models/flat_prior_mod.bug")

```

### Run Model

```{r, fig.width=7}
jags.data <- list(
  y = poor_sample$volunteer,
  x1 = poor_sample$precariat,
  x2 = poor_sample$lower_class,
  x3 = poor_sample$communal_lower_class,
  x4 = poor_sample$emergent_middle_class,
  x5 = poor_sample$private_middle_class,
  x6 = poor_sample$higher_middle_class,
  x7 = poor_sample$established_upper_class,
  N = nrow(poor_sample)
)

fit_flat_prior <- jags.model(
  file = "models/flat_prior_mod.bug",
  data = jags.data, 
  n.chains = 4
)

flat_results <- coda.samples(fit_flat_prior, 
                      variable.names = c("alpha", "beta1", "beta2", 
                                         "beta3", "beta4", "beta5", 
                                         "beta6", "beta7"),
                      n.iter = 10000, thin = 5)


## Save Model
save_it(flat_results)

summary(flat_results)



```

### Model Diagnostics

Some functions to create ggplot graphs

```{r}
no_plot <- function(model = NULL) {
  png("temp.xyz")
  plot_df <- gelman.plot(model)
  dev.off()
  file.remove("temp.xyz")
  return(invisible(plot_df))
}

gelman_plot <- function(codas, ...) {
  
  gp.dat <- no_plot(codas)
  
    df = data.frame(
      bind_rows(as.data.frame(gp.dat[["shrink"]][,,1]), 
                as.data.frame(gp.dat[["shrink"]][,,2])), 
                q=rep(dimnames(gp.dat[["shrink"]])[[3]], each=nrow(gp.dat[["shrink"]][,,1])),
                last.iter=rep(gp.dat[["last.iter"]], length(gp.dat))
      )
  
  reshape2::melt(df, c("q","last.iter"), value.name = "shrink_factor") %>% 
    ggplot(aes(last.iter, shrink_factor, colour = q, linetype = q)) + 
    geom_hline(yintercept=1, colour="grey30", lwd=0.2) +
    geom_line() +
    facet_wrap(~variable, labeller= labeller(.cols=function(x) gsub("V", "Chain ", x)), ...) +
    labs(x="Last Iteration in Chain", 
         y="Shrink Factor",
       colour="Quantile", 
       linetype="Quantile") +
  scale_linetype_manual(values = c(2,1))
}
```

Actual diagnostics:

```{r, echo = T, fig.width=12, fig.height=10}
tidy_flat_results <- ggs(flat_results) 

f1 <- ggs_traceplot(tidy_flat_results)  +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Trace Diagnostic Plot - Flat Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15)) 

f2 <- ggs_density(tidy_flat_results)  +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Density Diagnostic Plot - Flat Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15))

grid.arrange(f1, f2, ncol = 2, nrow = 1)

ggs_autocorrelation(tidy_flat_results) +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Autocorrelation Diagnostic Plot - Flat Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15)) +
  facet_grid(Parameter ~ Chain, labeller= labeller(.cols=function(x) paste0("Chain ", x)))


gelman_plot(flat_results, nrow = 2)   +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Gelman Diagnostic Plot - Flat Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15))
```


### Model Results

```{r, echo = T, fig.width=12, fig.height=10}
tidy_bayes_dat <- flat_results %>% 
  tibble(output = .) %>% 
  mutate(chain = as.character(1:n())) %>%
  mutate(output = map(output, as_tibble)) %>% 
  tidyr::unnest() %>% 
  group_by(chain) %>% 
  mutate(step = 1:n()) %>% 
  ungroup

expo <- function(x) {
  exp(x)/(1+exp(x))
}

intercept_dat <- tidy_bayes_dat %>% 
  select(alpha)

ridges_dat_flat <- tidy_bayes_dat %>% 
  select(-chain, -step) %>% 
  mutate_at(vars(beta1:beta7), function(x) intercept_dat$alpha + x) %>% 
  mutate_all(expo) %>% 
  # summarise_all(mean)
  gather(key, value) %>% 
  mutate(key = case_when(
    key == "beta1" ~ "Precariat (beta1)",
    key == "beta2" ~ "Lower class (beta2)",
    key == "beta3" ~ "Communal lower class (beta3)",
    key == "beta4" ~ "Emergent middle class (beta4)",
    key == "beta5" ~ "Private middle class (beta5)",
    key == "beta6" ~ "Higher middle class (beta6)",
    key == "beta7" ~ "Established upper class (beta7)",
    key == "alpha" ~ "Secure middle class (alpha)"
  )) %>% 
  mutate(key = fct_relevel(key, c("Precariat (beta1)",
                                  "Lower class (beta2)",
                                  "Communal lower class (beta3)",
                                  "Emergent middle class (beta4)",
                                  "Private middle class (beta5)",
                                  "Secure middle class (alpha)",
                                  "Higher middle class (beta6)",
                                  "Established upper class (beta7)"
                                  )))

# c("Established upper class", 
# "Higher middle class", "Secure middle class", "Private middle class", 
# "Emergent middle class", "Communal lower class", "Lower class", "Precariat",
# )
  
mean_dat_flat <- ridges_dat_flat %>% 
  group_by(key) %>% 
  dplyr::summarize_all(mean)

ridges_dat_flat %>%
  ggplot(aes(x = value, y = key)) +
  geom_halfeyeh() +
  stat_intervalh(aes(x = value), .width = c(.5, .8, .95)) +
  # geom_point(aes(x = mean(value), group = key)) +
  scale_color_grey("High Density Intervals", start = 0.5, end = 0.1) +
  theme_hc() +
  geom_text(data = mean_dat_flat,
            aes(x = value, label = round(value, 2), group = key), 
            nudge_y = 0.2) +
  geom_point(data = mean_dat_flat, aes(x = value, group = key), color = "white") +
  labs(title = "Predicted Probablities for each Class - Flat Prior", 
       x = "Predicted Probabilities",
       y = "") +
  theme(legend.position = "bottom", 
        plot.title = element_text(size = 25),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))


```

A different way to quickly visualize logit coefficients:

```{r}
ggs_caterpillar(tidy_flat_results)
```


## Compare to Rich Sample


```{r, echo = T, fig.width=12, fig.height=10}

glm_dat <- get_model_data(glm_mod, type = "pred", show.intercept = T) %>% 
  magrittr::extract2(1) %>% as_tibble() %>% 
  mutate(x = case_when(
    x == 1 ~ "Secure middle class (alpha)",
    x == 8 ~ "Precariat (beta1)",
    x == 7 ~ "Lower class (beta2)",
    x == 6 ~ "Communal lower class (beta3)",
    x == 5 ~ "Emergent middle class (beta4)",
    x == 4 ~ "Private middle class (beta5)",
    x == 3 ~ "Higher middle class (beta6)",
    x == 2 ~ "Established upper class (beta7)"
  )) %>% 
  mutate(x = fct_relevel(x, c("Precariat (beta1)",
                                  "Lower class (beta2)",
                                  "Communal lower class (beta3)",
                                  "Emergent middle class (beta4)",
                                  "Private middle class (beta5)",
                                  "Secure middle class (alpha)",
                                  "Higher middle class (beta6)",
                                  "Established upper class (beta7)"
                                  )))

ridges_dat_flat %>%
  ggplot(aes(x = value, y = key)) +
  geom_halfeyeh() +
  stat_intervalh(aes(x = value), .width = c(.5, .8, .95)) +
  # geom_point(aes(x = mean(value), group = key)) +
  scale_color_grey("High Density Intervals", start = 0.5, end = 0.1) +
  theme_hc() +
  geom_text(data = mean_dat_flat,
            aes(x = value, label = round(value, 2), group = key), 
            nudge_y = 0.2) +
  geom_point(data = mean_dat_flat, aes(x = value, group = key), color = "white") +
  labs(title = "Predicted Probablities for each Class - Flat Prior vs. Gezellig et al.", 
       x = "Predicted Probabilities",
       y = "",
       subtitle = "Estimates in Red are from Gezellig et al.") +
  theme(legend.position = "bottom", 
        plot.title = element_text(size = 20),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))  +
  geom_errorbarh(data = glm_dat,
                 aes(x = predicted, y = x,
                     xmin = conf.low, 
                     xmax = conf.high,
                     height = 0), inherit.aes = F, 
                 position = position_nudge(y = 0.4), color = "red")  +
  geom_point(data = glm_dat,
                 aes(x = predicted, y = x), 
                   inherit.aes = F, 
             position = position_nudge(y = 0.4), color = "red")  +
  geom_text(data = glm_dat,
                 aes(x = predicted, y = x, label = round(predicted, 2)), 
                   inherit.aes = F, 
             position = position_nudge(y = 0.55, x = 0.01), color = "red") 


```

## Informative Priors

```{r}
prior_dat <- glm_mod %>% 
  broom::tidy() %>% 
  mutate(term = str_remove_all(term, "social_class")) %>% 
  mutate(term = factor(term, levels = c("(Intercept)", "Precariat",
                                        "Lower class",
                                        "Communal lower class",
                                        "Emergent middle class", 
                                        "Private middle class",
                                        "Higher middle class", 
                                        "Established upper class"))) %>% 
  arrange(term) %>% 
  mutate(inverse_variance = 1/(std.error^2))

prior_dat
```


### Model Specification

```{r binary_model, echo = T}

info_prior_mod <- "model{
  for (i in 1:N){
    y[i] ~ dbern(p[i])
    logit(p[i]) <- alpha + beta1 * x1[i] + beta2 * x2[i]+ beta3 * x3[i] + beta4 * x4[i]+ beta5 * x5[i] + beta6 * x6[i] + beta7 * x7[i] 
  }
  alpha ~ dnorm(mu_alpha, var_alpha)
  beta1 ~ dnorm(mu_beta1, var_beta1)
  beta2 ~ dnorm(mu_beta2, var_beta2)
  beta3 ~ dnorm(mu_beta3, var_beta3)
  beta4 ~ dnorm(mu_beta4, var_beta4)
  beta5 ~ dnorm(mu_beta5, var_beta5)
  beta6 ~ dnorm(mu_beta6, var_beta6)
  beta7 ~ dnorm(mu_beta7, var_beta7) 
}"

write(info_prior_mod, "models/info_prior_mod.bug")

```

### Run Model

```{r, fig.width=15}
jags.data <- list(
  y = poor_sample$volunteer,
  x1 = poor_sample$precariat,
  x2 = poor_sample$lower_class,
  x3 = poor_sample$communal_lower_class,
  x4 = poor_sample$emergent_middle_class,
  x5 = poor_sample$private_middle_class,
  x6 = poor_sample$higher_middle_class,
  x7 = poor_sample$established_upper_class,
  mu_alpha = prior_dat$estimate[1],
  mu_beta1 = prior_dat$estimate[2],
  mu_beta2 = prior_dat$estimate[3],
  mu_beta3 = prior_dat$estimate[4],
  mu_beta4 = prior_dat$estimate[5],
  mu_beta5 = prior_dat$estimate[6],
  mu_beta6 = prior_dat$estimate[7],
  mu_beta7 = prior_dat$estimate[8],
  var_alpha = prior_dat$inverse_variance[1],
  var_beta1 = prior_dat$inverse_variance[2],
  var_beta2 = prior_dat$inverse_variance[3],
  var_beta3 = prior_dat$inverse_variance[4],
  var_beta4 = prior_dat$inverse_variance[5],
  var_beta5 = prior_dat$inverse_variance[6],
  var_beta6 = prior_dat$inverse_variance[7],
  var_beta7 = prior_dat$inverse_variance[8],
  N = nrow(poor_sample)
)

fit_info_prior <- jags.model(
  file = "models/info_prior_mod.bug",
  # inits = jags.inits,
  data = jags.data, 
  n.chains = 4
)

info_results <- coda.samples(fit_info_prior, 
                      variable.names = c("alpha", "beta1", "beta2", 
                                         "beta3", "beta4", "beta5", 
                                         "beta6", "beta7"),
                      n.iter = 10000, thin = 5)



save_it(info_results)

summary(info_results)

```

### Model Diagnostics


```{r, echo = T, fig.width=12, fig.height=10}
tidy_info_results <- ggs(info_results) 

f1 <- ggs_traceplot(tidy_info_results)  +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Trace Diagnostic Plot - Informative Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15)) 

f2 <- ggs_density(tidy_info_results)  +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Density Diagnostic Plot - Informative Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15))

grid.arrange(f1, f2, ncol = 2, nrow = 1)

ggs_autocorrelation(tidy_info_results) +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Autocorrelation Diagnostic Plot - Informative Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15)) +
  facet_grid(Parameter ~ Chain, labeller= labeller(.cols=function(x) paste0("Chain ", x)))


gelman_plot(info_results, nrow = 2)   +
  theme_hc() +
  scale_color_gdocs() +
  scale_fill_gdocs() +
  labs(title = "Gelman Diagnostic Plot - Informative Prior\n") +
  theme(legend.position = "none", 
        plot.title = element_text(size = 25),
        strip.text = element_text(size = 15))
```

### Model Results


```{r, echo = T, fig.width=12, fig.height=10}
tidy_bayes_dat <- info_results %>% 
  tibble(output = .) %>% 
  mutate(chain = as.character(1:n())) %>%
  mutate(output = map(output, as_tibble)) %>% 
  tidyr::unnest() %>% 
  group_by(chain) %>% 
  mutate(step = 1:n()) %>% 
  ungroup


intercept_dat <- tidy_bayes_dat %>% 
  select(alpha)

ridges_dat_info <- tidy_bayes_dat %>% 
  select(-chain, -step) %>% 
  mutate_at(vars(beta1:beta7), function(x) intercept_dat$alpha + x) %>% 
  mutate_all(expo) %>% 
  # summarise_all(mean)
  gather(key, value) %>% 
  mutate(key = case_when(
    key == "beta1" ~ "Precariat (beta1)",
    key == "beta2" ~ "Lower class (beta2)",
    key == "beta3" ~ "Communal lower class (beta3)",
    key == "beta4" ~ "Emergent middle class (beta4)",
    key == "beta5" ~ "Private middle class (beta5)",
    key == "beta6" ~ "Higher middle class (beta6)",
    key == "beta7" ~ "Established upper class (beta7)",
    key == "alpha" ~ "Secure middle class (alpha)"
  )) %>% 
  mutate(key = fct_relevel(key, c("Precariat (beta1)",
                                  "Lower class (beta2)",
                                  "Communal lower class (beta3)",
                                  "Emergent middle class (beta4)",
                                  "Private middle class (beta5)",
                                  "Secure middle class (alpha)",
                                  "Higher middle class (beta6)",
                                  "Established upper class (beta7)"
                                  )))
  
mean_dat_info <- ridges_dat_info %>% 
  group_by(key) %>% 
  dplyr::summarize_all(mean)

ridges_dat_info %>%
  ggplot(aes(x = value, y = key)) +
  geom_halfeyeh() +
  stat_intervalh(aes(x = value), .width = c(.5, .8, .95)) +
  # geom_point(aes(x = mean(value), group = key)) +
  scale_color_grey("High Density Intervals", start = 0.5, end = 0.1) +
  theme_hc() +
  geom_text(data = mean_dat_info,
            aes(x = value, label = round(value, 2), group = key), 
            nudge_y = 0.2) +
  geom_point(data = mean_dat_info, aes(x = value, group = key), color = "white") +
  labs(title = "Predicted Probablities for each Class - Informative Prior", 
       x = "Predicted Probabilities",
       y = "") +
  theme(legend.position = "bottom", 
        plot.title = element_text(size = 25),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))

```


## Compare all of Them

```{r, echo = T, fig.width=12, fig.height=10}

compare_dat <- bind_rows(
  ridges_dat_flat %>% 
    mutate(type = "Flat Prior"),
  
  ridges_dat_info %>% 
    mutate(type = "Informative Prior")
)

compare_dat_mean <- bind_rows(
  mean_dat_flat %>% 
    mutate(type = "Flat Prior"),
  
  mean_dat_info %>% 
    mutate(type = "Informative Prior")
)

compare_dat %>%
  ggplot(aes(x = value, y = key, fill = type), alpha = 0.02) +
  # geom_density(alpha = 0.5) +
  geom_halfeyeh() +
  # facet_wrap(~key, nrow = 2) +
  # stat_intervalh(aes(x = value), .width = c(.5, .8, .95)) +
  scale_fill_gdocs("Model"
                   #, start = 0.5, end = 0.1
                   ) +
  theme_hc() +
  ggrepel::geom_text_repel(data = compare_dat_mean,
            aes(x = value, label = round(value, 3), group = key),
            nudge_y = 0.2) +
  geom_point(data = compare_dat_mean, aes(x = value, group = key),
             color = "white",  show.legend = F) +
  labs(title = "Predicted Probablities for each Class - Informative Prior",
       x = "Predicted Probabilities",
       y = "") +
  theme(legend.position = "bottom", 
        plot.title = element_text(size = 25),
        axis.text = element_text(size = 15),
        axis.title = element_text(size = 15))  +
  geom_errorbarh(data = glm_dat,
                 aes(x = predicted, y = x,
                     xmin = conf.low, 
                     xmax = conf.high,
                     height = 0), inherit.aes = F, 
                 position = position_nudge(y = 0.4), color = "darkgreen")  +
  geom_point(data = glm_dat,
                 aes(x = predicted, y = x), 
                   inherit.aes = F,  show.legend = F,
             position = position_nudge(y = 0.4), color = "darkgreen")  +
  geom_text(data = glm_dat,
                 aes(x = predicted, y = x, label = round(predicted, 2)), 
                   inherit.aes = F, 
             position = position_nudge(y = 0.55, x = 0.01), color = "darkgreen") 

```